I am most grateful to my grandparents, Dimas and Sileide, who were present at all stages of this important moment in my trajectory.
I am grateful not only for all the sacrifices made to provide me with a solid education, but also for all the support and encouragement to overcome the various challenges that have surfaced over the past two years.
You are my greatest inspiration in life.

I would also like to offer a special thanks to my fiancee, Juliana.
Thank you for being part of my life and especially for being by my side during this important moment.
Her company, love and patience were essential in this work.
Her smile was and is my biggest motivation.
Thanks for everything.

I also take the opportunity to register my sincere thanks to my advisor Prof. Dr. Heitor Ramos.
I am extremely grateful for your orientation, advice and all the opportunities offered.
I feel that thanks to you I have evolved as a student and researcher in those years.

I would also like to thank my co-advisor and mentor Prof. Dr. Alejandro Frery.
I feel extremely lucky after leaving each of our conversations and meetings, where even from a distance, I feel motivated with his love for science.
I deeply hope that our current partnership is only the beginning of a great history of friendship.

To my friends, Glauber Braga, Bruno Rafael and David Alexandre, who endured all my crises and outbursts and were present in the best and worst moments of the last few years.
Love you!

To my new "brother friend", Pedro Barros, who embarked on this journey with me from my hometown, Maceió.
Thank you for all of our valued conversations and scientific debates present at Laccan House every day.

I would like to thank the professors André Aquino, André Lages, Osvaldo Rosso and Eliana Almeida for having believed on my academic potential.
The support provided by you was crucial.
Thank you!

Finally, and no less important, I would like to thank the examining board for their attentive reading, questions and suggestions.